---
title: "675_midterm"
author: "Gillian Xuezhu Zhao"
date: "3/17/2021"
output: word_document
---
# Load libraries and setup
```{r setup}
library(dplyr)
library(tidyverse)
library(sf)
library(RSocrata)
library(viridis)
library(caret)
library(spatstat)
library(spdep)
library(FNN)
library(grid)
library(gridExtra)
library(knitr)
library(kableExtra)
library(plotROC)
library(ROCR)
library(pROC)
library(raster)
library(stargazer)


# Gillian's working directory
#setwd("G:\\UPenn\\CPLN675_LandUse_and_EnvironmentalModeling\\midterm\\CPLN-675-Midterm")

# Xinyi's working directory
setwd("/Users/xinyi/Desktop/Courses/Land Use and Environmental Modeling/Midterm/task 4 Building Model")

# Nearest neighbor function
nn_function <- function(measureFrom,measureTo,k) {
  measureFrom_Matrix <-
    as.matrix(measureFrom)
  measureTo_Matrix <-
    as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
    output <-
      as.data.frame(nn) %>%
      rownames_to_column(var = "thisPoint") %>%
      gather(points, point_distance, V1:ncol(.)) %>%
      arrange(as.numeric(thisPoint)) %>%
      group_by(thisPoint) %>%
      summarize(pointDistance = mean(point_distance)) %>%
      arrange(as.numeric(thisPoint)) %>% 
      dplyr::select(-thisPoint) %>%
      pull()
  
  return(output)  
}

# Map theme for maps
mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 14,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    plot.background = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
  )
}

# Color for ggplots
## Calgary
c_colors_hex <- c("#4a6fa5", "#dbe9ee")
names(c_colors_hex) <- c("land", "stream")
## Denver
d_colors_hex <- c("#4f6d7a", "#c0d6df")
names(d_colors_hex) <- c("land", "stream")

```

# Calgary
## Load in features (from task 1 and task 2)
```{r additional feature manipulation}
c_landcover_net <-st_read ('c_landcover_net/c_landcover_net.shp')
c_ndvi_net <- st_read ('c_ndvi_net/c_ndvi_net.shp')
c_vars_net <- st_read("c_vars_net.shp")
c_vars_net <- cbind(c_vars_net,c_landcover_net,c_ndvi_net) %>%
  dplyr::select(elevatn, 
                slope, gntlSlp, stepSlp, 
                gntlS_1, gntlS_2, gntlS_3, 
                stpSl_1, stpSl_2, stpSl_3,
                mainBsn, 
                dwnst_1, dwnst_2, dwnst_3,resist,ndvi)
#st_write(c_vars_net, "c_vars_net_update.shp")

c_vars_net <- st_read ('c_vars_net_update/c_vars_net_update.shp')

```

## Visualize features in fishnet grid
```{r fishnet feature visualization}
c_vars <- 
  c_vars_net %>%
  dplyr::select(elevatn, 
                slope, gntlSlp, stepSlp, 
                gntlS_1, gntlS_2, gntlS_3, 
                stpSl_1, stpSl_2, stpSl_3,
                mainBsn, 
                dwnst_1, dwnst_2, dwnst_3,resist,ndvi)

c_vars_long <- 
  c_vars %>%
  gather(Variable, value, -geometry)

c_vars <- unique(c_vars_long$Variable)
c_mapList <- list()

for(i in c_vars){
  c_mapList[[i]] <- 
    ggplot() +
    geom_sf(data = filter(c_vars_long, Variable == i), aes(fill=value), colour=NA) +
    scale_fill_viridis(name="") +
    labs(title=i) +
    mapTheme()}

do.call(grid.arrange,c(c_mapList, ncol =4, top = "Risk Factors by Fishnet"))
```

# Denver
## Load in features (from task 1)
```{r additional feature manipulation}
d_landcover_net <-st_read('d_landcover_net/d_landcover_net.shp')
d_ndvi_net <- st_read('d_ndvi_net/d_ndvi_net.shp')
d_vars_net <- st_read("d_vars_net.shp")
d_vars_net <- cbind(d_vars_net,d_landcover_net,d_ndvi_net) %>%
  dplyr::select(elevatn, 
                slope, gntlSlp, stepSlp, 
                gntlS_1, gntlS_2, gntlS_3, 
                stpSl_1, stpSl_2, stpSl_3,
                mainBsn, 
                dwnst_1, dwnst_2, dwnst_3,resist,d_ndv_r)%>%
  rename(ndvi=d_ndv_r)
#st_write(d_vars_net, "d_vars_net_update.shp")

d_vars_net <- st_read ('d_vars_net_update/d_vars_net_update.shp')
```

## Visualize features in fishnet grid
```{r fishnet feature visualization}
d_vars <- 
  d_vars_net %>%
  dplyr::select(elevatn, 
                slope, gntlSlp, stepSlp, 
                gntlS_1, gntlS_2, gntlS_3, 
                stpSl_1, stpSl_2, stpSl_3,
                mainBsn, 
                dwnst_1, dwnst_2, dwnst_3,resist,ndvi)

d_vars_long <- 
  d_vars %>%
  gather(Variable, value, -geometry)

d_vars <- unique(d_vars_long$Variable)
d_mapList <- list()

for(i in d_vars){
  d_mapList[[i]] <- 
    ggplot() +
    geom_sf(data = filter(d_vars_long, Variable == i), aes(fill=value), colour=NA) +
    scale_fill_viridis(name="") +
    labs(title=i) +
    mapTheme()}

do.call(grid.arrange,c(d_mapList, ncol =4, top = "Risk Factors by Fishnet"))
```


# Calgary Flood Inundation
```{r}
c_flood<-raster('c_flood.tif')
c_boundary <- 
  st_read("calgary_boundary/CALGIS_CITYBOUND_LIMIT.shp") %>%
  st_transform(crs=3780)
c_floodp <-
  rasterToPoints(c_flood) %>%
  as.data.frame() %>%
  st_as_sf (coords=c('x','y'),crs= st_crs(c_flood))%>%
  st_transform(st_crs(c_boundary))


```

```{r}
c_fishnet <- 
  st_make_grid(c_boundary, cellsize = 500) %>%
  st_sf() %>%
  st_transform(crs=3780)

c_fishnet <- 
  c_fishnet[c_boundary,] %>%
  mutate(uniqueID = rownames(.)) %>%
  dplyr::select(uniqueID)
```

```{r}
c_flood_net<-
  c_floodp %>%
  dplyr::filter(c_flood ==1) %>%
  aggregate(.,c_fishnet, mean) %>%
  mutate(c_flood = ifelse(is.na(c_flood), 0, c_flood))

#sf::st_write(c_flood_net,'c_flood_net.shp')
  
```

```{r}
ggplot()+
  geom_sf(data=c_flood_net ,aes (fill=as.factor(c_flood)),color=NA) +
  scale_fill_manual (values = c('grey','dark red'),
                     labels = c('No inundation','Inundation'))+
  labs(title= 'Calgary Flood Inundation')+
  mapTheme()+
  theme(legend.title = element_blank())
  
  
```

# Create Variables Dataset
```{r}
model_var_net <-cbind(c_vars_net,c_flood_net)%>%
  dplyr::select(elevatn, 
                slope, gntlSlp, stepSlp, 
                gntlS_1, gntlS_2, gntlS_3, 
                stpSl_1, stpSl_2, stpSl_3,
                mainBsn, 
                dwnst_1, dwnst_2, dwnst_3,resist,ndvi,c_flood) %>%
  rename (flood = c_flood)

model_var_net$mainBsn <- as.factor(model_var_net$mainBsn)
  
```
## Standardize
```{r}
c_final <-
  model_var_net %>%
  filter(is.na(elevatn)==FALSE & is.na(slope)==FALSE)%>%
  mutate(elevation = (elevatn-min(elevatn))/(max(elevatn)-min(elevatn))) %>%
  mutate(slopee = (slope-min(slope))/(max(slope)-min(slope))) %>%
  mutate(gntlslope = (gntlSlp-min(gntlSlp))/(max(gntlSlp)-min(gntlSlp))) %>%
  mutate(stepslope = (stepSlp-min(stepSlp))/(max(stepSlp)-min(stepSlp))) %>%
  dplyr::select(-elevatn,-slope,-gntlSlp,-stepSlp)
  
```

# Build Model
```{r}
set.seed(3456)
trainIndex <-createDataPartition(c_final$mainBsn , p=.70 ,
                                  list=FALSE,
                                  times=1)
modelTrain<-c_final[trainIndex,]
modelTest <-c_final[-trainIndex,]
```

```{r}
floodModel <- glm(flood ~.,family = "binomial"(link="logit"),
                  data= modelTrain %>%
                    as.data.frame() %>%
                    dplyr::select(-geometry))
summary(floodModel)
```

## Exclude insignificant variables
```{r}
c_final2 <-
  c_final %>%
  dplyr::select (gntlS_3,dwnst_3 ,resist,ndvi,elevation,gntlslope,mainBsn,flood)

modelTrain<-c_final2[trainIndex,]
modelTest <-c_final2[-trainIndex,]

floodModel <- glm(flood ~.,family = "binomial"(link="logit"),
                  data= modelTrain %>%
                    as.data.frame() %>%
                    dplyr::select(-geometry))
summary(floodModel)
step(floodModel)
```

## Final Model
```{r}
c_final3 <-
  c_final %>%
  dplyr::select (gntlS_3,resist,ndvi,elevation,gntlslope,mainBsn,flood)

modelTrain<-c_final3[trainIndex,]
modelTest <-c_final3[-trainIndex,]

floodModel <- glm(flood ~.,family = "binomial"(link="logit"),
                  data= modelTrain %>%
                    as.data.frame() %>%
                    dplyr::select(-geometry))
summary(floodModel)
stargazer(floodModel,title="Flood Prediction Model Results",align=TRUE,type='text')
```


### Visualization Four Significant Features
```{r}
c_sig_vars <-
  c_vars_net %>%
  dplyr::select(elevatn,mainBsn,resist,ndvi)

c_sig_vars_long <-
  c_sig_vars %>%
  gather(Variable,value,-geometry)

c_sig_vars <- unique(c_sig_vars_long$Variable)
c_sig_mapList <-list()
  
for(i in c_sig_vars){
  c_sig_mapList[[i]] <- 
    ggplot() +
    geom_sf(data = filter(c_sig_vars_long, Variable == i), aes(fill=value), colour=NA) +
    scale_fill_viridis(name="") +
    labs(title=i) +
    mapTheme()}

do.call(grid.arrange,c(c_sig_mapList, ncol =2, top = "Risk Factors by Fishnet"))
```
```{r}
c_plot_vars<-
  model_var_net %>%
  as.data.frame() %>%
  dplyr::select(flood,elevatn,gntlSlp,resist,ndvi,gntlS_3) %>%
  gather(key,value,elevatn:gntlS_3) %>%
  mutate(value= ifelse(key=="resist",value*1000,
                ifelse(key=="ndvi", value*5000,
                ifelse(key=="gntlSlp",value*10,value))))

ggplot(c_plot_vars, aes(as.factor(flood),value, fill=as.factor(flood)))+
  geom_bar(stat="identity")+
  facet_wrap(~key)+
  scale_fill_manual(values=c("grey","darkred"),
                    labels = c("No Inundation","Inundation"),
                    name="")+
  labs(x="Flood",y="Value")
```
```{r}
c_plot_Basin<-
  model_var_net %>%
  as.data.frame() %>%
  dplyr::select(flood,mainBsn) %>%
  group_by(flood,mainBsn) %>%
  summarise(count=n())

ggplot(c_plot_Basin, aes(x=as.factor(flood), y=count,fill=as.factor(mainBsn)))+
  geom_bar(stat="identity",position ="stack")+
  scale_fill_manual(values=c("grey","darkred"),
                    labels = c("Not Main Watershed","Main Watershed"),
                    name="")+
  labs(x="Flood",y="Count")
```

## Model Validation
```{r}
classProbs <- predict (floodModel , modelTest,type='response')

testProbs <- data.frame (obs = as.numeric(modelTest$flood),pred = classProbs)

ggplot(testProbs, aes(x=pred, fill = as.factor(obs)))+
  geom_density()+
  facet_grid(obs~.) + xlab('Probability') + geom_vline(xintercept = .5) +
  scale_fill_manual(values = c("grey", "dark red"),
                      labels = c("No Inundation","Inundation"),
                      name = "")


```

#### Since the model is designed to predict the flood inundation, we hope the sensitivity ( or recall) of the model could be as high as possible so that we cannot miss any flood inundation. While, at the meantime, since government needs to invest in plenty of money, labors and time to improve the facilities where has larger possibilities of flooding, we hope the percesion, which is a ratio of correctly predicted positive observations to the total predicted positive observations, could be as high as possible either. To balance the recall and percesion, I choose F-1 score to find the best threshold. And the threshold is 0.432.
```{r}

pred<-prediction(testProbs[is.na(testProbs$pred)==FALSE,]$pred,testProbs[is.na(testProbs$pred)==FALSE,]$obs)
f.perf<-performance(pred,"f")
plot(f.perf)

F.score <-c(f.perf@y.values[[1]])
cutoff<-c(f.perf@x.values[[1]])
F.score_table<-data.frame(cbind(F.score,cutoff))
F.score_table[which.max(F.score_table$F.score),]
```



```{r}
testProbs$predClass  = ifelse(testProbs$pred > .438 ,1,0)

caret::confusionMatrix(reference = as.factor(testProbs$obs), 
                       data = as.factor(testProbs$predClass), 
                       positive = "1")
```






```{r}
ggplot(testProbs, aes(d = obs, m = pred)) + 
  geom_roc(n.cuts = 50, labels = FALSE) + 
  style_roc(theme = theme_grey) +
  geom_abline(slope = 1, intercept = 0, size = 1.5, color = 'grey') 

auc(testProbs$obs, testProbs$pred)
```

## k-folds Cross Validation: use entire dataset
```{r}
ctrl <- trainControl(method = "cv", 
                     number = 100, 
                     classProbs=TRUE,
                     savePredictions = TRUE,summaryFunction=twoClassSummary)

cvFit <- train(as.factor(flood) ~ .,  data = c_final3 %>% 
                                             as.data.frame() %>%
                                             dplyr::select(-geometry)%>%
                                             mutate(flood = ifelse(flood==1,'yes','no')), 
               method="glm", family="binomial",
               metric="ROC",
               trControl = ctrl,
               na.action=na.exclude)

cvFit

```

```{r}
dplyr::select(cvFit$resample, -Resample) %>%
  gather(metric, value) %>%
  left_join(gather(cvFit$results[2:4], metric, mean)) %>%
  ggplot(aes(value)) + 
    geom_histogram(bins=35, fill = "#FF006A") +
    facet_wrap(~metric) +
    geom_vline(aes(xintercept = mean), colour = "#981FAC", linetype = 3, size = 1.5) +
    scale_x_continuous(limits = c(0, 1)) +
    labs(x="Goodness of Fit", y="Count", title="CV Goodness of Fit Metrics for Flood Inundation Model",
         subtitle = "Across-fold mean reprented as dotted lines")

cvFit$results
```

# Calgary Flood Predictions
```{r}
allPredictions<-
  predict(cvFit, c_final3,type="prob")[,2]

c_final3<-
  cbind(c_final3,allPredictions)%>%
  mutate(allPredictions=round(allPredictions*100))
```

```{r}
 ggplot() + 
    geom_sf(data=c_final3, aes(fill=factor(ntile(allPredictions,5))), colour=NA) +
    scale_fill_manual(values = c("#edf8fb","#b3cde3","#8c96c6","#8856a7","#810f7c"),
                      labels=as.character(quantile(c_final3$allPredictions,
                                                 c(0.1,.2,.4,.6,.8),na.rm=T)),
                      name="Predicted\nProbabilities(%)\n(Quintile\nBreaks)") +
  mapTheme() +
  labs(title="Predicted Probabilities of Flood Inundation in Calgary")
```

```{r}
ggplot() + 
  geom_sf(data=c_final3, aes(fill=factor(ntile(allPredictions,5))), colour=NA) +
  scale_fill_manual(values = c("#edf8fb","#b3cde3","#8c96c6","#8856a7","#810f7c"),
                    labels=as.character(quantile(c_final3$allPredictions,
                                                 c(0.1,.2,.4,.6,.8),na.rm=T)),
                    name="Predicted\nProbabilities(%)\n(Quintile\nBreaks)") +
  geom_sf(data=c_flood_net %>% 
               filter(c_flood == 1), 
               fill="red",colour=NA) +
  mapTheme() +
  labs(title="Predicted Probabilities and Past Flood",
       subtitle="Calgary: Past flood in red")
```

## Prediction Accuracy Map for training set
```{r}
trainProbs <-predict(cvFit,modelTrain,type="prob")[,2]

Train_confusionmatrix<-
  cbind(modelTrain,trainProbs) %>%
  mutate(trainProbs = round(trainProbs*100)) %>%
  mutate(predict_flood = ifelse (trainProbs > 43.8,1,0)) %>%
  mutate(confusion_matrix =
           ifelse(predict_flood==1 & flood ==1, "True_Positive",
           ifelse(predict_flood==1 & flood ==0,"False_Positive",
           ifelse(predict_flood==0 & flood ==0, "True_Negative",
           ifelse(predict_flood==0 & flood ==1, "False_Negative",0)))))
ggplot()+
  geom_sf(data=Train_confusionmatrix, aes(fill=factor(confusion_matrix)),color=NA) +
  scale_fill_manual(values=c("red","lightskyblue","grey","royalblue"),
                    labels=c("False Negative","False Positive","True Negative","True Positive"),
                    name = "Confusion Matrix for Traning Set")+
  mapTheme()


```

## Inundation Map for Calgary
```{r}
c_predict_flood<-
  c_final3 %>%
  mutate(predict_flood = ifelse(allPredictions>43.8,1,0))

ggplot()+
  geom_sf(data=c_predict_flood, aes(fill=factor(predict_flood)),color=NA)+
  scale_fill_manual(values = c("grey","darkred"),
                    labels=c("No Inundation","Inundation"),
                    name="Flood Prediction")+
  mapTheme()
```

# Prediction for Denver
## Visualization
```{r}
d_sig_vars <-
  d_vars_net %>%
  dplyr::select(elevatn,mainBsn,resist,ndvi)

d_sig_vars_long <-
  d_sig_vars %>%
  gather(Variable,value,-geometry)

d_sig_vars <- unique(d_sig_vars_long$Variable)
d_sig_mapList <-list()
  
for(i in d_sig_vars){
  d_sig_mapList[[i]] <- 
    ggplot() +
    geom_sf(data = filter(d_sig_vars_long, Variable == i), aes(fill=value), colour=NA) +
    scale_fill_viridis(name="") +
    labs(title=i) +
    mapTheme()}

do.call(grid.arrange,c(d_sig_mapList, ncol =2, top = "Risk Factors by Fishnet"))
```

## Prediction
```{r}
D.model_var_net <-
  d_vars_net%>%
  dplyr::select(gntlS_3,resist,ndvi,elevatn,gntlSlp,mainBsn)%>%
  mutate(elevation = (elevatn-min(elevatn))/(max(elevatn)-min(elevatn))) %>%
  mutate(gntlslope = (gntlSlp-min(gntlSlp))/(max(gntlSlp)-min(gntlSlp))) %>%
  dplyr::select(-elevatn,-gntlSlp)
  

D.model_var_net$mainBsn<-as.factor(D.model_var_net$mainBsn)

DProb <-predict(floodModel,D.model_var_net,type="response")

Denver_flood<-
  cbind(D.model_var_net,DProb)%>%
  mutate(allPredictions = round(DProb * 100)) %>%
  mutate(allPredictions=ifelse(is.na(allPredictions)==FALSE,allPredictions,0))

ggplot() + 
    geom_sf(data=Denver_flood, aes(fill=factor(ntile(allPredictions,5))), colour=NA) +
    scale_fill_manual(values = c("#edf8fb","#b3cde3","#8c96c6","#8856a7","#810f7c"),
                      labels=as.character(quantile(Denver_flood$allPredictions,
                                                 c(0.1,.2,.4,.6,.8),na.rm=T)),
                      name="Predicted\nProbabilities(%)\n(Quintile\nBreaks)") +
  mapTheme() +
  labs(title="")
```


```{r}
d_predict_flood<-
  Denver_flood %>%
  mutate(predict_flood = ifelse(allPredictions>43.8,1,0))

ggplot()+
  geom_sf(data=d_predict_flood, aes(fill=factor(predict_flood)),color=NA)+
  scale_fill_manual(values = c("grey","darkred"),
                    labels=c("No Inundation","Inundation"),
                    name="Flood Prediction")+
  mapTheme()
```

